---
title: "Recurrent Neural Network"
author: "Bastiaan Quast"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Recurrent Neural Network}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Package

This package includes an example Recurrent Neural Network.
The package is loaded using:

```{r package}
library(rnn)
```

# Code

We can view the code of the main `rnn()` function by calling it without the parathesis (not printed here).

```{r code-rnn, eval=FALSE}
trainr
```

As can be seen from the above, the model relies on two other functions that are available through the `sigmoid` package.

The first function is `logistic()`, which converts an integer to its sigmoid value.

```{r sigmoid}
(a <- sigmoid::logistic(3))
```

The code for the `sigmoid()` function is:

```{r sigmoid-code}
sigmoid::logistic
```

The second function converts the sigmoid value of a number to its derivative.

```{r sigmoid-der}
sigmoid::sigmoid_output_to_derivative(a) # a was created above using sigmoid()
```

Finally, we can inspect this code using:

```{r sigmoid-der-code}
sigmoid::sigmoid_output_to_derivative
```

# Application

By setting a **seed** for the random number generator, we ensure replicability.

```{r seed}
set.seed(1)
```

An example is included in the help file.

```{r help, eval=FALSE}
help('trainr')
```

Below is a basic function that converts integers to binary format (read left to right)

```{r int2bin}
# basic conversion
i2b <- function(integer, length=8)
  as.numeric(intToBits(integer))[1:length]

# apply to entire vectors
int2bin <- function(integer, length=8)
  t(sapply(integer, i2b, length=length))
```

First we generate the data:

```{r data}
# create sample inputs
X1 = sample(0:127, 5000, replace=TRUE)
X2 = sample(0:127, 5000, replace=TRUE)

# create sample output
Y <- X1 + X2

# convert to binary
X1 <- int2bin(X1)
X2 <- int2bin(X2)
Y  <- int2bin(Y)

# Create 3d array: dim 1: samples; dim 2: time; dim 3: variables.
X <- array( c(X1,X2), dim=c(dim(X1),2) )
Y <- array( Y, dim=c(dim(Y),1) ) 
```

This example is:

```{r example}
# train the model
model <- trainr(Y=Y[,dim(Y)[2]:1,,drop=F], # we inverse the time dimension
                X=X[,dim(X)[2]:1,,drop=F], # we inverse the time dimension
                learningrate   =  0.1,
                hidden_dim     = 10,
                batch_size = 100,
                numepochs = 10)
```

See the evolution of the error over different epochs:

```{r error}
plot(colMeans(model$error),type='l',
     xlab='epoch',
     ylab='errors'                  )
```

Now create testing data

```{r test-data}
# create test inputs
A1 = int2bin( sample(0:127, 7000, replace=TRUE) )
A2 = int2bin( sample(0:127, 7000, replace=TRUE) )

# create 3d array: dim 1: samples; dim 2: time; dim 3: variables
A <- array( c(A1,A2), dim=c(dim(A1),2) )
```

Predict based on testing data.

```{r predictr}
# predict
B  <- predictr(model,
               A[,dim(A)[2]:1,,drop=F]
               )
B = B[,dim(B)[2]:1]
```

Define basic functions to convert binary to integer

```{r bin2int}
b2i <- function(binary)
  packBits(as.raw(c(binary, rep(0, 32-length(binary) ))), 'integer')

bin2int <- function(binary){
  binary <- round(binary)
  length <- dim(binary)[2]    # determine length of binary representation
  apply(binary, 1, b2i)     } # apply to full matrix
```

Test prediction against true values

```{r test}
# convert back to integers
A1 <- bin2int(A1)
A2 <- bin2int(A2)
B  <- bin2int(B)

# plot the difference
hist( B-(A1+A2) )
```
